### 2024-07-09 cognitive-bias.qmd
20240708 - regression fallacy

>>>>

The regression fallacy is the failure to account for the fact that things tend to naturally fluctuate around their mean and that, after an extreme fluctuation, subsequent scores tend to regress (or reverse) to the mean.
An example of the regression fallacy is the so-called Sports Illustrated or Madden cover jinx curse.
The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated or Madden will perform poorly.
But, such a phenomenon can be more simply explained by regression to the mean (<https://www.psychologytoday.com/us/blog/what-the-luck/201610/the-sports-illustrated-cover-jinx>; archived at <https://perma.cc/CZM9-TVFN>).
When a player has a superb season, they likely benefited from some degree to good luck, and it is unlikely that they will repeat such a stellar season the following year.
Instead, they are likely—at least based on random fluctuation—to regress to their long-term mean.

Applied to fantasy football, consider that a Quarterback had a 5-touchdown game in Week 1.
You are in need of a strong Quarterback, so you drop a solid player to pick him up.
However, it is possible that the Quarterback benefited from playing against a week defense in the first game of the season.
Future matchups may prove more difficult, and the player is unlikely to sustain such a solid performance consistently throughout the season.
